\documentclass[]{homework}

\begin{document}


\homework{10}{April 16, 2021}

\begin{problem}{1}
  On the previous homework, we looked at a zero-padding transformation in Fourier space.
  Here we will investigate the impact of re-sampling.

  % Clarify convention: no 1/N

  As with last time, consider a signal $h(t_n)$ sampled at N values, and a corresponding Fourier transform $H(\nu_m)$.
  \begin{subproblem}{a}
    {\em Up-sampling}: Suppose we construct a new signal from $H(\nu_m)$ by doubling the number of points as follows.
    Let $m' = 0, 1, ...\,, 2N-1$ and
    \[
      G(\nu_{m'}) = \begin{cases} 
      H(\nu_{m'/2}), & m'\,{\rm even} \\
      0 & m'\,{\rm odd}.
      \end{cases}
    \]
    Show that
    \[
      g(t_{n'}) = \begin{cases} 
      h(t_{n'}), & 0 \le n' < N \\
      h(t_{n'-N}), & N \le n' < 2N\,.
   \end{cases}
    \]
  \end{subproblem}
  \begin{subproblem}{b}
    {\em Down-sampling}: Suppose we construct a new signal from $H(\nu_m)$ by removing the frequencies with odd $m$,
    that is for $m' = 0, 1, ...\,, N/2-1$ we let $G(\nu_{m'}) = H(\nu_{2m'})$. Show that
    \[
      g(t_{n'}) = \frac{1}{2}\left( h(t_{n'}) + h(t_{n'+N/2}) \right)\,.
    \]
  \end{subproblem}
\end{problem}

\begin{problem}{2}
  In place of a final exam, we will have a final project.
  Your goal for this week will be to develop a project topic.
  Make sure to choose something you think you can complete with a reasonable
  amount of work--this should be roughly 3 homework assignments worth of effort.
  To get a sense for the scope of a project idea, here is a list of project
  topics from a similar course:\\
  \url{https://www.asc.ohio-state.edu/physics/ntg/780/project_list.php}

  \begin{subproblem}{a}
    Read through the project guidelines (available on Canvas).
  \end{subproblem}
  \begin{subproblem}{b}
    Provide a short written description of your chosen topic, a few sentences and/or bullet points.
    Outline what steps you will need to take to complete the project.
  \end{subproblem}
  \begin{subproblem}{c}
      Sign up for an individual meeting time on Friday to discuss your chosen topic,\\
      \url{https://www.signupgenius.com/go/904044EA5A82EA5FE3-p427}
  \end{subproblem}
\end{problem}

\vspace{1cm}

{\bf Optional midterm review:} Review your exam, and correct any incorrect or missing work.
For a fully corrected exam, you can collect up to half of the grade you missed (eg. $C \rightarrow B$).

\end{document}

