\documentclass[]{homework}

\begin{document}


\homework{7}{March 18, 2021}

\begin{problem}{1} 
  We saw in class it was possible to derive a variety of finite-difference formulae for
  computing derivatives of arbitrary order.

  \begin{subproblem}{a}
    On the midterm, we derived an expression for centered differencing that was accurate
    to third order, $\mathcal{O}(\Delta x^2)$. A higher-order expression, valid to order
    $\mathcal{O}(\Delta x^4)$, is given by
    \[
      f'(x) = \frac{-f(x+2\Delta x) + 8f(x+\Delta x) - 8f(x-\Delta x) + f(x-2\Delta x)}{12 \Delta x}\, + \mathcal{O}(\Delta x^4).
    \]
    Verify this is correct by Taylor expanding the various $f$ on the right-hand
    side of this expression and canceling terms.
  \end{subproblem}

  \begin{subproblem}{b}
    We can similarly derive a formula accurate to $\mathcal{O}(\Delta x^2)$
    for second derivatives. As above, verify that
    \[
      f''(x) = \frac{f(x+\Delta x) - 2f(x) + f(x-\Delta x)}{\Delta x^2}\,+ \mathcal{O}(\Delta x^2).
    \]
  \end{subproblem}

\end{problem}


\begin{problem}{2}
  In class, we looked at the PDE\_Integration notebook for solving the
  advection equation. This notebook can be downloaded from Canvas.
  \begin{subproblem}{a}
    Modify the notebook to solve the diffusion equation,
    \[
      \partial_t f(x, t) = c\, \partial_x^2 f(x, t)\,.
    \]
    Use the second-order finite-difference formula derived in problem 1b.
    You can use the same numerical coefficient c,
    $\Delta x$, and $\Delta t$ in the notebook, although feel free
    to experiment with changing these. You should find that the initially
    Gaussian profile ``diffuses'', becoming a broader, shorter distribution.
  \end{subproblem}
  \begin{subproblem}{b}
    It is very difficult to write down stable solutions for this system when
    $c$ is negative. Verify that your solution will not
    provide a valid answer if you integrate for a sufficiently long time with
    a negative value of $c$.
    Can you think of a physical reason why this system is unstable?
  \end{subproblem}
\end{problem}


\end{document}

